---
title: "705372204_stats102a_hw2"
author: "Hashim Bhat (705372204)"
date: "2023-01-25"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)

source("~/705372204_stats102a_hw2.R")
knitr::opts_chunk$set(echo = TRUE)

```

## Question 1

(a)

Create a simulated dataset in R1 called gradebook that represents a possible gradebook in
the basic format as Table 1:

```{r}
set.seed(999999999)
gradebook <- data.frame(UID = runif(100, 111111111, 10^9) %/% 1,
                        Homework_1 = 50 + runif(100, 0, 50) %/% 1,
                        Homework_2 = 50 + runif(100, 0, 50) %/% 1,
                        Homework_3 = 50 + runif(100, 0, 50) %/% 1,
                        Homework_4 = 50 + runif(100, 0, 50) %/% 1,
                        Homework_5 = 50 + runif(100, 0, 50) %/% 1,
                        Quiz_1 = 50 + runif(100, 0, 50) %/% 1,
                        Quiz_2 = 50 + runif(100, 0, 50) %/% 1,
                        Quiz_3 = 50 + runif(100, 0, 50) %/% 1,
                        Quiz_4 = 50 + runif(100, 0, 50) %/% 1,
                        Quiz_5 = 50 + runif(100, 0, 50) %/% 1)
head(gradebook)
```

(b)

Write R code in R markdown file to randomly replace 10% of Homework_4 and Quiz_4 by
NA respectively, and then use is.na() in conjunction with sum() to show your results.

```{r}
gradebook$Homework_4[sample(100,10)] <- NA
gradebook$Quiz_4[sample(100,10)] <- NA
sum(is.na(gradebook))
```

(c)

Write a function `messy_impute()` that imputes missing values in the gradebook. Please
also present your algorithm or flowchart to answer this question in the R markdown file.

#### **`messy_impute()`**

**Algorithm:**


1. Standardize center input to lowercase using tolower
2. Make sure first input is a dataframe by using if statement, handle error at the end of entire function
3. If center = mean, assign a mean function with variable name "center_function"
4. If center = median, assign a median function with same variable name "center_function"
5. Handle error for second input using stop function, specifying that input needs to be 'mean' or 'median'
6. For third input, use if statements for if margin is 1 and if margin is 2
7. If margin is 1, use for loop to apply center function to NA values, imputing them with the respective column averages
8. If margin is 2, use for loops and grepl functions to impute NA values with respective row averages, separating Homework and Quiz columns so that the row averages are calculated based on the assignment type. 
9. Handle error for third input using stop function, specifying that input must be either 1 or 2
10. Return the updatd dataframe at the end of the function


(d)


**Test Cases:**
```{r}
test_1 <- messy_impute(gradebook, "mean", 1)
head(test_1)

test_2 <- messy_impute(gradebook, "median", 2)
head(test_2)

test_3 <- messy_impute(gradebook, "mean", 1, trim = 0.25)
head(test_3)
```


(e)

Write R code using the main function in the tidyr package to convert the gradebook into
the tidy format. Name this object gradebook_tidy. You may directly write code in the R
markdown file. 

```{r}
gradebook_tidy <- as_tibble(gradebook) %>%
                  pivot_longer(names(gradebook)[-1], names_to = "Assignment", values_to = "Score") %>%
                  separate(Assignment, into = c("Assignment_Type", "Assignment_Number"))

head(gradebook_tidy)
```


(f)

Write a function tidy_impute() that imputes missing values in gradebook_tidy object.
The tidy_impute() function should have the same arguments as in the messy_impute()
function. You should return an imputed gradebook_tidy object. Please also present your
algorithm or flowchart to answer this question in the R markdown file.

#### **`tidy_impute()`**

**Algorithm:**

1. Standardize second input to lower case using tolower function
2. If center = mean, assign a mean function with variable name "center_fn"
3. If center = median, assign a median function with same variable name "center_fn"
3. Handle error for second input using stop function, specifying that input needs to be 'mean' or 'median'
4. If margin = 1, group the data by UID and assignment number using group_by function
5. If margin = 2, group the data by UID and assignment type using group_by function
6. Handle error for margin using stop function, specifying that input needs to be either 1 or 2
7. Use mutate function in combination with if_else function and center_fn to replace the NA values in Score column with average valuse calculated by center_fn. If condition is met (condition being that the value is NA), center_fn is applied and if condition is not met, data remains the same. 
8. Return the updated dataframe. 


(g)

```{r}

test_4 <- tidy_impute(gradebook_tidy, 'mean', 2)
head(test_4)

test_5 <- tidy_impute(gradebook_tidy, "median", 2)
head(test_5)

test_6 <- tidy_impute(gradebook_tidy, "mean", 1, trim = 0.25)
head(test_6)

```

## Question 2

(a)

**First Example**

One example of non-tidy data is Retail Sales Data such as the example provided below that shows historical sales data for 45 stores located in different regions where each store contains a number of departments. The company also runs several promotional markdown events throughout the year. The variables include Store Number, Date, Temperature in the region, cost of fuel in the region, markdown data related to promotional events, CPI, unemployment rate, isHoliday (which denotes whether the week is a special holiday week.)

```{r}
Features_data_set <- read.csv("~/Documents/Features_data_set.csv")
head(Features_data_set)
```

Dataset can be found here:
https://www.kaggle.com/datasets/manjeetsingh/retaildataset?resource=download

**Second Example**

Second example is Food Insecurity Data like the example below. The variables display peoples situation as it relates to food insecurity. Other variables are age, Sex, Race, Education. I removed some columns from the original to make it simpler. 

```{r}
Final_Pulse_Data <- read.csv("~/Documents/Final_Pulse_Data.csv")
Food_Insecurity <- Final_Pulse_Data[-c(10:40)]
head(Food_Insecurity)
```

Dataset can be found here:
https://www.kaggle.com/datasets/jackogozaly/pulse-survey-food-insecurity-data

**Third Example**

Billboard data that displays observations about tracks, displaying variables like artist, year, genre, and billboard ranking in certain weeks. 

Dataset can be found here:
https://github.com/hadley/tidy-data/blob/master/data/billboard.csv


```{r}
billboard <- read.csv("~/Documents/billboard.csv")
ncol(billboard)
billboard_simple <- billboard[-c(12:83)]
head(billboard_simple)
```


(b)

**First Example**

For the Sales Data, it could be presented better by having a column that denotes markdown number and another column that shows the corresponding markdown data. We can do this using as_tibble and pivot_longer functions.

```{r}
sales_tidy <- as_tibble(Features_data_set) %>%
              pivot_longer(names(Features_data_set)[5:9], names_to = "MarkDown_Number", values_to = "Markdown") 

sales_tidy
```

**Second Example**

Instead of having different column for food status, we could make one column that displays food status and then add a frequency count column.

```{r}
Food_ins_tidy <- as_tibble(Food_Insecurity) %>%
              pivot_longer(names(Food_Insecurity)[1:5], names_to = "Food_Status", values_to = "Frequency_Count") 

Food_ins_tidy
```


**Third Example**

There are multiple columns for week. We should instead have one single column that displays week number, and then add a column that displays the billboard ranking in that particular week. We can do this using as_tibble and pivot_longer functions.

```{r}
billboard_tidy <- as_tibble(billboard_simple) %>%
                  pivot_longer(names(billboard_simple)[8:11], names_to = "Week_Number", values_to = "Billboard_Ranking")
billboard_tidy
```



